Simultaneous Elicitation of Scoring Rule and Agent Preferences for Robust Winner Determination

### 2021-05-11 minimax.tex
PMR and MR formulas in just one line

Given $\pprofile$ and $\pw$ (that represent the current knowledge about agent preferences and the PSR),
the maximum regret is considered by assuming that an adversary can both 1) extend the partial profile $\pprofile$ into a complete profile, and 2) instantiate the weights choosing among any weight vector in $\pw$. %where $\pprofile$ and $\pw$ represent our knowledge so far.
We formalize the notion of minimax regret in multiple steps.
First of all, $\Regret(x, \profile, \w)$
is the “regret” of selecting $x$ as a winner instead of the optimal alternative under $\profile$ and $\w$:
$$\Regret(x, \profile, \w) = \max_{y \in A} s(y; \profile,\w) - s(x; \profile, \w).$$
The {\em pairwise maximum regret} of $x$ relative to $y$ given the partial profile $\pprofile$ and the set of weights $W$ is the worst-case loss of choosing $x$ instead of $y$ under all possible realizations of the full profile {\em and} all possible instantiations of the weights:
$$\PMR(x,y;\pprofile,W) = \max_{\w \in W} \max_{\profile \in C(\pprofile)} s(y; \profile,\w) - s(x; \profile,\w).$$
The maximum regret %$\MaxR(x;\pprofile,W)$ 
is the worst-case loss of $x$:
\begin{align}
	\MaxR(x;\pprofile,W) & = \max_{y \in A} \PMR(x,y; \pprofile, W) \\ &= \max_{\w \in W} \max_{\profile \in C(\profile)} \Regret(x; \profile, \w).
\end{align}
$\MaxR(x;\pprofile,W)$ is the result of an adversarial selection of the complete profile $\profile \in C(\pprofile)$ and of the scoring vector $\w \in W$ that jointly maximize the loss between $x$ and the true winner under $\profile$ and $\w$.
Finally,  $\MMR(\pprofile,W) = \min_{x \in A} \MaxR(x;\pprofile,W)$ is the value of {\em minimax regret} under $\pprofile$ and $W$, obtained when recommending a {\em minimax optimal} alternative $x^*_{\pprofile, W} \in A^*_{\pprofile, W} = \argmin_{x \in A} \MaxR(x;\pprofile,W)$.
Picking as consensus choice an alternative associated with minimax regret provides a recommendation that gives worst-case guarantees. 
In cases of ties, %in minimax regret, 
we can return all minimax alternatives $A^*_{\pprofile, W}$ as winners or pick one of them using some tie-breaking strategy.

Observe that if $\MMR(\pprofile, W)\!=\!0$, then any $x^{*}_{\pprofile,W} \in A^*_{\pprofile, W}$ is a necessary winner: any valid completion of the profile and choice of $w \in W$ gives to $x^{*}_{\pprofile,W}$ the highest score.

@@

Note that in \cref{claim:rankPMR}, in the case $(x \pprefeq_j y)$, $\beta$ is the number of alternatives incomparable with both $x$ and $y$.
\begin{proposition}\label{claim:PMR}
	The $\PMR$ can be written as:
	\begin{align} 
		\PMR(x,y; \pprofile, W)  
		& = \max_{w \in W} \sum_{j \in N} w_{\hat{v}_j(y)} - w_{\hat{v}_j(x)} \\ 
		& = \max_{w \in W} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_i 
	\end{align}
	where $\hat{\alpha}_{r}^{y}$ (resp. $\hat{\alpha}_{r}^{x}$)  is the number of times $y$ (resp. $x$) has rank $r$ in the complete profile $\hat{\profile}$ defined in \cref{claim:rankPMR}. 
\end{proposition}
%\cref{claim:rankPMR} can be understood by observing that in the case $(x \pprefeq_j y)$, $\beta$ is the number of alternatives incomparable with both $x$ and $y$.
\cref{claim:PMR}  shows that PMR is linear in the weights.
%Recall that $\pw$ represents the weight vectors compatible with the preferences of the chair stated so far.
The pairwise max regret $\PMR(x,y; \pprofile,W)$ can thus be obtained by solving the following linear program defined on the variables $w_1, …, w_m$:
\begin{align}
	\max_{\w} & \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_{r}\\
	\text{ s.t. } & w_1 = 1 ≥ … ≥ w_m = 0, \text{\cref{eq:convexity}} \text{ and } \w \in \pw.
\end{align}
%Note that given our choice $w_{1}=1$ and $w_{m}=0$, there are only $m-2$ variables 
%(we leave $w_{1}$ and $w_{m}$ in the LP just for clarity of presentation).
The max regret $\MaxR(x; \pprofile, W)$ is determined by computing the pairwise regret of $x$ with all other alternatives in $A$, and the recommended alternatives are the ones with least max regret. 
Observe that when the $\PMR$ of an alternative $x$ (against some other alternative $y$) exceeds the best MR value found so far, we do not need to further evaluate $x$. 
This idea can be exploited using a minimax-search tree \cite{Braziunas2012}.

>>>>

Given $\pprofile$ and $\pw$ (that represent the current knowledge about agent preferences and the PSR),
the maximum regret is considered by assuming that an adversary can both 1) extend the partial profile $\pprofile$ into a complete profile, and 2) instantiate the weights choosing among any weight vector in $\pw$. %where $\pprofile$ and $\pw$ represent our knowledge so far.
We formalize the notion of minimax regret in multiple steps.
First of all, $\Regret(x, \profile, \w)$
is the “regret” of selecting $x$ as a winner instead of the optimal alternative under $\profile$ and $\w$:
$$\Regret(x, \profile, \w) = \max_{y \in A} s(y; \profile,\w) - s(x; \profile, \w).$$
The {\em pairwise maximum regret} of $x$ relative to $y$ given the partial profile $\pprofile$ and the set of weights $W$ is the worst-case loss of choosing $x$ instead of $y$ under all possible realizations of the full profile {\em and} all possible instantiations of the weights:
$$\PMR(x,y;\pprofile,W) = \max_{\w \in W} \max_{\profile \in C(\pprofile)} s(y; \profile,\w) - s(x; \profile,\w).$$
The maximum regret %$\MaxR(x;\pprofile,W)$ 
is the worst-case loss of $x$:
\begin{align}
	\MaxR(x;\pprofile,W) & = \max_{y \in A} \PMR(x,y; \pprofile, W) %\\ &
	= \max_{\w \in W} \max_{\profile \in C(\profile)} \Regret(x; \profile, \w).
\end{align}
$\MaxR(x;\pprofile,W)$ is the result of an adversarial selection of the complete profile $\profile \in C(\pprofile)$ and of the scoring vector $\w \in W$ that jointly maximize the loss between $x$ and the true winner under $\profile$ and $\w$.
Finally,  $\MMR(\pprofile,W) = \min_{x \in A} \MaxR(x;\pprofile,W)$ is the value of {\em minimax regret} under $\pprofile$ and $W$, obtained when recommending a {\em minimax optimal} alternative $x^*_{\pprofile, W} \in A^*_{\pprofile, W} = \argmin_{x \in A} \MaxR(x;\pprofile,W)$.
Picking as consensus choice an alternative associated with minimax regret provides a recommendation that gives worst-case guarantees. 
In cases of ties, %in minimax regret, 
we can return all minimax alternatives $A^*_{\pprofile, W}$ as winners or pick one of them using some tie-breaking strategy.

Observe that if $\MMR(\pprofile, W)\!=\!0$, then any $x^{*}_{\pprofile,W} \in A^*_{\pprofile, W}$ is a necessary winner: any valid completion of $\pprofile$ and any choice of $w \in W$ gives to $x^{*}_{\pprofile,W}$ the highest score.

@@

Note that in \cref{claim:rankPMR}, in the case $(x \pprefeq_j y)$, $\beta$ is the number of alternatives incomparable with both $x$ and $y$.
\begin{proposition}\label{claim:PMR}
	The $\PMR$ can be written as:
	\begin{align} 
		\PMR(x,y; \pprofile, W)  
		& = \max_{w \in W} \sum_{j \in N} w_{\hat{v}_j(y)} - w_{\hat{v}_j(x)} %\\ &
		 = \max_{w \in W} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_i 
	\end{align}
	where $\hat{\alpha}_{r}^{y}$ (resp. $\hat{\alpha}_{r}^{x}$)  is the number of times $y$ (resp. $x$) has rank $r$ in the complete profile $\hat{\profile}$ defined in \cref{claim:rankPMR}. 
\end{proposition}
%\cref{claim:rankPMR} can be understood by observing that in the case $(x \pprefeq_j y)$, $\beta$ is the number of alternatives incomparable with both $x$ and $y$.
\cref{claim:PMR}  shows that PMR is linear in the weights.
%Recall that $\pw$ represents the weight vectors compatible with the preferences of the chair stated so far.
The pairwise max regret $\PMR(x,y; \pprofile,W)$ can thus be obtained by solving the following linear program defined on the variables $w_1, …, w_m$:
\begin{align}
	\max_{\w} & \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_{r}\\
	\text{ s.t. } & w_1 = 1 ≥ … ≥ w_m = 0, \text{\cref{eq:convexity}} \text{ and } \w \in \pw.
\end{align}
%Note that given our choice $w_{1}=1$ and $w_{m}=0$, there are only $m-2$ variables 
%(we leave $w_{1}$ and $w_{m}$ in the LP just for clarity of presentation).
The max regret $\MaxR(x; \pprofile, W)$ is determined by computing the pairwise regret of $x$ with all other alternatives in $A$, and the recommended alternatives are the ones with least max regret. 
Observe that when the $\PMR$ of an alternative $x$ (against some other alternative $y$) exceeds the best MR value found so far, we do not need to further evaluate $x$. 
This idea can be exploited using a minimax-search tree \cite{Braziunas2012}.
